---
title: "R Notebook"
output: html_notebook
---

This notebook is for Feature Selection.


Step 3: feature selection
```{r}
library(Boruta)

ED.final<-read.csv("~/GitHub/Spr2017-proj5-grp3/data/NAreplaced.csv")
#names(ED.final[,1])="challengeID"
label<-read.csv("~/Documents/FFChallenge/train.csv")


label<-label[,1:2]
label=na.omit(label)
Index=ED.final$challengeID %in% label$challengeID

data.train<-ED.final[Index,]
data.train<-as.data.frame(data.train)

model.features = Boruta(data.train[,-1], label.train$gpa, maxRuns = 500)

results = as.data.frame(model.features$finalDecision)
which(results!="Rejected")
Bcodes = rownames(results)[which(results!="Rejected")]

```


What are the descriptions of the features (Boruta)?
```{r}
des = data.frame()
#score = data.frame()
for (i in 1:length(Bcodes))
{
  if(Bcodes[i] %in% data.info$code)
  {
    des = rbind(des, as.data.frame(data.info$description[data.info$code==Bcodes[i]]))
   #score = rbind(score, as.data.frame(imp$meanImp[which(rownames(imp)==Bcodes[i])]))

  }

}

decision = as.data.frame(model.features$finalDecision[model.features$finalDecision!="Rejected"])

Bdf = cbind(Bcodes, des, decision)
colnames(Bdf) = c("Codes", "Description", "Decision")

pred="gpa"
write.csv(Bdf, paste0("../data/",pred,"features.csv"))
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
